
```{r message=FALSE}
library(data.table)
library(stringr)
library(ggplot2)

# Load utility functions
source('~/bayes-rate-consistency/R/covimod-utility.R')

# Set ggplot theme
theme_set(theme_bw())
```

```{r message=FALSE}
# Load the COVIMOD data
covimod <- load_covimod_data("~/bayes-rate-consistency")
dt.part <- covimod$part
dt.hh <- covimod$hh
dt.nhh <- covimod$nhh
dt.pop <- covimod$pop

# Load Germany district data
dt.district <- as.data.table(readr::read_csv('~/bayes-rate-consistency/data/germany-districts.csv'))
```

## Preprocessing
```{r}
## Participant data
# Remove participants with missing age-strata or gender information
dt.part <- dt.part[!is.na(gender) & !is.na(age_strata)]

# Impute children age by sampling from uniform distribution
dt.part <- impute_child_age(dt.part, seed=1527)

# Urban / Rural classification
dt.part <- merge(dt.part, dt.district[, .(district, form)], 
                 all.x = TRUE, 
                 by = "district",
                 allow.cartesian = TRUE)
setindex(dt.part, new_id, wave, date)
setcolorder(dt.part, c("new_id", "wave", "date", "age", "age_strata", "gender", "district", "form"))

dt.pop <- age_stratify(dt.pop)
```


## Sample size analysis
### Number of survey participants by residence
```{r}
dt.part[wave == 1, .(N = .N), by = .(form)]
dt.part[wave == 1, .(N = .N), by = .(form, imp_age)]
```
### Number of observed contacts by residence and type
```{r}
tmp <- dt.hh[wave == 1 & hh_met_this_day == 1]
tmp <- merge(tmp, dt.part, by = c("new_id", "wave"), all.x = TRUE)
su.hh <- tmp[, .(N = .N), by = .(form)]
su.hh[, type := "household"]

tmp <- dt.nhh[wave == 1]
tmp <- merge(tmp, dt.part, by = c("new_id", "wave"), all.x = TRUE)
su.nhh <- tmp[, .(N = .N), by = .(form)]
su.nhh[, type := "non-household"]

su <- rbind(su.hh, su.nhh)
setkey(su, form, type)
setcolorder(su, c("form", "type", "N"))

su[, prop := N / sum(N), by = form]
su
```

## Distribution of contacts
### Average number of contacts by region
Calculate the number of household contacts urban / rural
```{r}
# Household
tmp <- dt.hh[wave == 1 & hh_met_this_day == 1]
tmp <- tmp[, .(contacts = .N), by = c("new_id", "wave")]
tmp <- merge(tmp, dt.part, all.x = TRUE, by = c("new_id", "wave"))
tmp.hh <- tmp[, .(new_id, contacts, age_strata, form)]

# Non household
tmp <- dt.nhh[wave == 1, .(contacts = .N), by = c("new_id", "wave")]
tmp <- merge(tmp, dt.part, all.x = TRUE, by = c("new_id", "wave"))
tmp.nhh <- tmp[, .(new_id, contacts, age_strata, form)]

tmp <- rbind(tmp.hh, tmp.nhh)
tmp <- tmp[, .(contacts = sum(contacts)), by = c("new_id", "age_strata", "form")]
tmp <- tmp[contacts < 30 & !is.na(age_strata)]

ggplot(tmp, aes(age_strata, contacts, color = form)) + 
  geom_jitter(alpha = 0.7) + 
  geom_boxplot(width = 0.6) + 
  ggsci::scale_color_d3() + 
  scale_y_continuous(limits = c(0, 25)) + 
  facet_wrap(~form) + 
  labs(x = "Age group", y = "Number of Contacts") + 
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    axis.text.x = element_text(angle = 45, hjust=1)
  )
```

```{r}
# Household
tmp <- dt.hh[wave == 1 & hh_met_this_day == 1]
tmp <- merge(tmp, dt.part, all.x = TRUE, by = c("new_id", "wave"))
tmp.hh <- tmp[, .(N = .N), by = .(form, imp_age, alter_age_strata)]

# Non household
tmp <- dt.nhh[wave == 1]
tmp <- merge(tmp, dt.part, all.x = TRUE, by = c("new_id", "wave"))
tmp.nhh <- tmp[, .(N = .N), by = .(form, imp_age, alter_age_strata)]

# Merge
tmp <- rbind(tmp.hh, tmp.nhh)
tmp <- tmp[, .(N_contact = sum(N)), by = .(form, imp_age, alter_age_strata)]
setnames(tmp, "imp_age", "age")

grid <- as.data.table(
  expand.grid(
    form = c("Urban", "Rural"),
    age = seq(0,84), 
    alter_age_strata = c("0-4","5-9","10-14","15-19","20-24","25-34","35-44",
                         "45-54","55-64","65-69","70-74","75-79","80-84")
  )
)

# Participant sample size
tmp.part <- dt.part[wave == 1, .(N_part = .N), by = .(form, imp_age)]
setnames(tmp.part, "imp_age", "age")

# Population size
tmp.pop <- dt.pop[, .(N_pop = sum(pop)), by = .(age_strata)]

# Combine everything
tmp <- merge(grid, tmp, all.x = TRUE)
tmp <- merge(tmp, tmp.part, all.x = TRUE, by = c("form", "age"))
tmp <- merge(tmp, tmp.pop, all.x = TRUE, by.x = "alter_age_strata", by.y = "age_strata")

tmp[N_part > 0 & is.na(N_contact), N_contact := 0]
tmp[, m := N_contact / N_part]
tmp <- tmp[!is.na(N_part)]
```

```{r}
ggplot(tmp, aes(age, alter_age_strata)) + 
  geom_tile(aes(fill = m)) + 
  viridis::scale_fill_viridis(option = "H") + 
  scale_x_continuous(expand = c(0,0), na.value = "white") + 
  scale_y_discrete(expand = c(0,0), na.value = "white") +
  labs(x = "Age of contacting individuals", y = "Age of contacts", fill = "Intensity") + 
  facet_wrap(~form) + 
  theme(
    aspect.ratio = 1,
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.margin = margin()
  )
```
## Bayes Rate Consistency
Compile Stan model
```{r}
library(cmdstanr)
model <- cmdstan_model("~/bayes-rate-consistency/stan_models/regional/hsgp-m52-rd.stan")
```

Configure Stan data
```{r}
source("~/bayes-rate-consistency/R/stan-utility.R")

stan_data <- list(
  N_U = nrow(tmp[form == "Urban"]),
  N_R = nrow(tmp[form == "Rural"]),
  A = 85,
  C = 13
)

# Add row major idx
tmp <- tmp[order(age, alter_age_strata, form)]
tmp[, age_idx := age + 1]
tmp[, age_strata_idx := as.numeric(alter_age_strata)]
tmp[, row_major_idx := (age_idx-1)*13 + age_strata_idx]

stan_data$ROW_MAJOR_IDX_U <- tmp[form == "Urban", row_major_idx]
stan_data$ROW_MAJOR_IDX_R <- tmp[form == "Rural", row_major_idx]

# Add contact counts
stan_data$Y_U <- tmp[form == "Urban", N_contact]
stan_data$Y_R <- tmp[form == "Rural", N_contact]

# Add offsets
grid <- as.data.table(expand.grid(age = seq(0,84), form = c("Urban", "Rural")))

tmp.offsets <- merge(grid, tmp.part, all.x = T)
tmp.offsets[is.na(N_part), N_part := 1]
stan_data$log_N_U <- log(tmp.offsets[form == "Urban", N_part])
stan_data$log_N_R <- log(tmp.offsets[form == "Rural", N_part])

stan_data$log_S_U <- log(rep(1, 85))
stan_data$log_S_R <- log(rep(1, 85))

stan_data$log_P <- log(dt.pop[age < 85, .(pop = sum(pop)), by=age]$pop)

stan_data <- add_map_age_to_strata(stan_data)
stan_data <- add_std_age_idx(stan_data)
stan_data <- add_nn_idx(stan_data)

stan_data <- add_hsgp_parms(stan_data, 1.5, 20, 30)
```

Set initial values
```{r}
beta_0 <- tmp[N_contact > 0, floor(mean( log( N_contact / N_part ) - log(N_pop) ))]
model_inits <- function(){ 
  list(
    beta_0 = beta_0 + rnorm(1, 0, 0.2),
    z = rnorm(1, 0, 0.01),
    nu = 0.4 + rnorm(1, 0, 0.01)
  ) 
}
```


Run the Stan model
```{r}
fit <- model$sample(
  data = stan_data, seed = 123, 
  chains = 2, iter_warmup = 500, iter_sampling = 500,
  parallel_chains = 2,
  init = model_inits,
  refresh = 25
)
```

Save the model
```{r}
fit$save_object("~/bayes-rate-consistency/stan_fits/hsgp-m52-rd-region.rds")
```

## Model diagnostics
```{r}
fit$diagnostic_summary()
fit$cmdstan_diagnose()
```

```{r}
fit$summary(variables = c("beta_0", "gp_alpha", "gp_rho_1", "gp_rho_2", "nu"))
```

Extract posterior contact intensities by residence type
```{r}
ps <- c(0.5, 0.025, 0.975)
p_labs <- c('M','CL','CU')

po <- fit$draws("log_cnt_rate", format = "draws_matrix")
dt.po <- as.data.table(reshape2::melt(po))

# Extract indices
.pattern <- "log_cnt_rate\\[([0-9]+),([0-9]+),([0-9]+)\\]"

dt.po[, region_idx := as.numeric(gsub(.pattern, "\\1", variable))]
dt.po[, age_idx := as.numeric(gsub(.pattern, "\\2", variable))]
dt.po[, alter_age_idx := as.numeric(gsub(.pattern, "\\3", variable))]

# Calculate quantiles
dt.po <- dt.po[, list(q=quantile(value, prob=ps, na.rm=T), q_label = p_labs),
                      by=list(region_idx, age_idx, alter_age_idx)]
dt.po <- data.table::dcast(dt.po, region_idx + age_idx + alter_age_idx ~ q_label, value.var = "q")

# Convert back to rates
dt.po[, M := exp(M)]
dt.po[, CL := exp(CL)]
dt.po[, CU := exp(CU)]

dt.po[, age := age_idx - 1]
dt.po[, alter_age := alter_age_idx - 1]
      
    # Recover gender and alter gender
dt.po[, form := ifelse(region_idx == 1, "Urban", "Rural")]

# Load datasets
dtp <- dt.pop[age < 85, .(pop = sum(pop)), by = age]
dt.po <- merge(dt.po, dtp, by.x="alter_age", by.y = "age", all.x = TRUE)

dt.po[, intensity_M := M * pop]
dt.po[, intensity_CL := CL * pop]
dt.po[, intensity_CU := CU * pop]
```

```{r}
ggplot(dt.po, aes(age, alter_age)) + 
  geom_tile(aes(fill = intensity_M)) + 
  coord_equal(expand=0) +
  viridis::scale_fill_viridis(option = "H") + 
  labs(x = "Age of contacting individuals", y = "Age of contacts", fill = "intensity") + 
  facet_wrap(~form) + 
  theme(
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.margin = margin()
  )
```

```{r}
extract_posterior_intensity <- function(po, dt.pop){
  po <- subset(po, "log_cnt_rate")
  dt.po <- as.data.table(reshape2::melt(po))

  # Extract indices
  .pattern <- "log_cnt_rate\\[([0-9]+),([0-9]+),([0-9]+)\\]"
  
  dt.po[, region_idx := as.numeric(gsub(.pattern, "\\1", variable))]
  dt.po[, age_idx := as.numeric(gsub(.pattern, "\\2", variable))]
  dt.po[, alter_age_idx := as.numeric(gsub(.pattern, "\\3", variable))]

  # Recover age and region
  dt.po[, age := age_idx - 1]
  dt.po[, alter_age := alter_age_idx - 1]
  dt.po[, form := ifelse(region_idx == 1, "Urban", "Rural")]

  # Remove unnecessary columns
  dt.po[, age_idx := NULL]
  dt.po[, alter_age_idx := NULL]
  dt.po[, region_idx := NULL]

  # Merge with population data
  dt.po <- merge(dt.po, dt.pop, by.x="alter_age", by.y="age", all.x = TRUE)

  # Calculate posterior contact intensities
  dt.po[, value := exp(value + log(pop))]

  return(dt.po)
}

dt.po <- extract_posterior_intensity(fit$draws("log_cnt_rate", format = "matrix"), dtp)
```

```{r}
dt <- dt.po[, .(value = sum(value)), by=c("draw", "age", "form")]
dt <- dt[, list( q=quantile(value, prob=ps, na.rm=T), q_label = p_labs), by=c("age", "form")]
dt <- data.table::dcast(dt, age + form ~ q_label, value.var = "q")

ggplot(dt, aes(age, M)) + 
  geom_line(aes(color = form), size = 0.7) + 
  geom_ribbon(aes(ymin = CL, ymax = CU, fill = form), alpha = 0.3) + 
  scale_x_continuous(expand = c(0,0)) + 
  scale_y_continuous(limits = c(0,10)) +
  ggsci::scale_color_lancet() + 
  ggsci::scale_fill_lancet() + 
  labs(x = "Age of contacting individuals", y = "Contact intensity", color="", fill="") + 
  theme(
    legend.position = "bottom",
    legend.margin = margin()
  )
```




